Bayesian exponential random graph modeling of KWL Burial network

### 2020-06-21 analysis/paper/paper.Rmd
edit method section

We analyzed archaeological burials from a Iron Age site, Kiwulan, in northeastern Taiwan to generate networks. We used burial data collected from the published excavation report, and original fieldwork notes with detailed descriptions about context and stratigraphy. We do not include the burials in Chinese phase in our network analysis because of the small sample size. <!-- but we will briefly discuss these in the discussion section --> We assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the quantity of burial goods. In addition to wealth, we also consider another two social variables, age and gender of the interred individual, serving as indicators for distinguishing ascribed or achieved status in a society [citation].

The purpose of our analysis is to understand the underlying mechanisms that shape the relationships between people at Kiwulan, and the effect of foreign contact on these relationships. We hypothesize that after the start of the European presence in the 17th century, a group of ambitious individuals competed with each other for access to trade with the Europeans due to the high value of trade goods in their society. This led to the accumulation of high quality trade goods among a small numbers of individuals. Social inequality gradually increased in Kiwulan, and we predict this can be observed from the burials. We hypothesize a change from a more equal distribution of grave goods to a more differentiated distribution of goods. Thus, we predict that the burial network will present less reciprocality, less transitivity, but higher popularity structure effects <!-- the reader needs to be introduced to these terms before you can use them like this... instead, need to foreground or foreshadow these concepts... "We used network analysis statistics to quantify the relationships and formally test our hypothesis. "--> in burial network after the European presence, compared to the network before European presence.

@@

We use Bayesian reference on Exponential Random Graph Models (ERGMs) to conduct statistical network analysis to test what variables may have contributed to the economic inequality observed between burials over time under foreign contact conditions.<!-- seems like a long sentence --> ERGMs assume that each network tie <!-- does the reader know what this is?--> is a random variable which is dependent on each other somehow through the shared attributes or a reciprocal relationship [@Robins2007]. The dependence of ties can be hypothesized by parameters corresponding to configurations in the network, such as a single tie, or a transitive triad <!-- does the reader know what is 'transitive? --> . A model is built up by a distribution of random graphs according to these configurations that are also called network statistics. According to the formation of ties, network statistics can be divided into two types, endogenous statistics and exogenous statistics [@Amati2019]. In endogenous models <!--statistic or model? what exactly? these seem like two different things that are getting mixed up. Choose one way to describe and stick to it to avoid confusing the reader -->, the dependence of ties is decided by the presence of other ties, while in an exogenous model, the dependence of ties is determined by node attributes <!-- node? -->. We use both exogenous and endogenous statistics to understand the characteristics of the burial network at Kiwulan and the correlate of burials. To describe the relationship linked by shared burial goods, the network is weighted to capture the multiple effects <!-- of what? the reader doesn't know about any effects yet --> [@Snijders2006]. The network variables we examine include connectivity, clusters, centralization, and homophily in undirected relations <!-- something wrong here, 'relations' seems wrong here --> to reflect the social relations <!-- avoid repeating the same work in a sentence --> between individuals guided by social structures. Table \@ref(tab:ergm-terms-table) lists the network statistics we used in the model construction for burials with archaeological interpretations <!-- put this earlier in the paragraph --> . Every term in an ERGM has an associated algorithm for computing its value for network [@Morris2008].

@@

The plots from left to right show estimated marginal posterior densities, trace plots, and autocorrelation plots.

@@

>>>>

We analyzed archaeological burials from a Iron Age site, Kiwulan, in northeastern Taiwan based on data collected from published excavation report, and original fieldwork notes to generate networks. The data includes information of gender, age, burial rites, and types of burial goods. We do not include the burials in Chinese phase in our network analysis because of the small sample size. A network is a structure consists of a set of nodes (actors) and edges (ties) connecting them representing relationships [@Brughmans2018]; in our case, burials are nodes linked by edges if any pair of burial has the prestige goods in common. This is based on the assumption that the prestige goods imply social practices of trade, exchange, and gifting that reflect social relations between individuals [@Coward2013, pp. 252]. Our purpose is to understand the underlying mechanisms that shape the relationships between people at Kiwulan, and the effect of foreign contact on these relationships. In other words, we explore which variables may have contributed to the observed economic inequality across burials using network analysis approaches, and any changes in network structure before and after the foreign contacts.

We hypothesized that after the start of the European presence in the 17th century, a group of ambitious individuals competed with each other for access to trade with the Europeans due to the high value of trade goods in their society. The high value of trade goods had been identified by the historical documents written by the Europeans and the Han Chinese [@Borao2009; @LiandWu2006]. In addition to valuable foreign goods, the indirect colonialism in northeastern Taiwan may suggest the potentially manipulation of the European colonial image by local ambitious individuals to built their personal status and power. This led to the monopolization of valuable goods among a small numbers of individuals and the society gradually switched to a structure with social differentiation. If the social inequality gradually increased in Kiwulan, then we will observe it from the burials by analyzing the networks generated by the social processes. This is important to understand how an Indigenous society reacted to the presence of foreign imperial power.

@@

We use Bayesian inference on exponential random graph models (ERGMs) to quantify the relationships among burials and test our hypothesis. ERGMs are statistical models for building complex social networks where possible ties are assumed as random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007]. A number of parameters used for modeling the dependencies expresses the structure or property of a network, such as reciprocity (relationship between two actors), transitivity (relationship between two actors through a shared third actor), homophily (relationship between actors with similar attribute), and popularity (actors have many relationships with others) [@Robins2007; @Snijders2006]. With a research question for a certain social process, we can build a hypothesis-based model to generate a distribution of random networks according to the parameters we specified, also called configurations or network statistics. This helps to understand the underlying processes of an observed network by comparing it with a large set of possible networks generated by the model.

Table \@ref(tab:ergm-terms-table) lists the parameters we used in the model construction for burials with archaeological interpretations. Every term in an ERGM has an associated algorithm for computing its value for network [@Morris2008]. Based on our hypothesis, the burial network after the European presence will present less reciprocity, less transitivity, but higher popularity structure effects compared to the network before the European presence. Since burials tend to have multiple types of burial goods in common, such as personal ornaments, imported porcelains, and gold foils, the network ties are weighted instead of a binary data expressed by 0 and 1 [@Snijders2006]. We assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the quantity of burial goods. In addition to wealth, we also consider another two social variables, age and gender of the interred individual, serving as indicators for distinguishing ascribed or achieved status in a society [citation].

@@

The plots from left to right show estimated marginal posterior densities, trace plots, and auto correlation plots.

@@

<!-- briefly discuss Chinese burials -->
